Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-14 fully-abstract-comp/technical-report.tex
removed opaque capabilities, added a small section about them, and added part of the unified semantics

% Text
\newcommand{\tand}{\text{ and }}
\newcommand{\totherwise}{\text{otherwise }}

@@

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{gettype}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
%\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
\newcommand{\sgetlin}[2]{\twoinstr{\src{get}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
\newcommand{\ssplice}[2]{\twoinstr{\src{splice}}{#1}{#2}}
\newcommand{\ssplit}[2]{\twoinstr{\src{split}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
\newcommand{\scseal}[3]{\threeinstr{\src{cseal}}{#1}{#2}{#3}}

@@

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{gettype}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
%\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
\newcommand{\tgetlin}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
\newcommand{\tsplice}[2]{\twoinstr{\trg{splice}}{#1}{#2}}
\newcommand{\tsplit}[2]{\twoinstr{\trg{split}}{#1}{#2}}

% Three arguments
\newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
\newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
\newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
\newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}
\newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

@@

%%% Permissions
\newcommand{\plainperm}[1]{\mathrm{#1}}
\newcommand{\rwlx}{\plainperm{rwlx}}
\newcommand{\rwx}{\plainperm{rwx}}
\newcommand{\rx}{\plainperm{rx}}
\newcommand{\rwlxo}{\plainperm{rwlxo}}
\newcommand{\rwlo}{\plainperm{rwlo}}
\newcommand{\rwl}{\plainperm{rwl}}
\newcommand{\rwxo}{\plainperm{rwxo}}
\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
\newcommand{\rxo}{\plainperm{rxo}}
\newcommand{\ro}{\plainperm{ro}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\noperm}{\plainperm{0}}
\newcommand{\nopermo}{\plainperm{0o}}
\newcommand{\enter}{\plainperm{e}}
\newcommand{\entero}{\plainperm{eo}}

@@

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePermission}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}

@@

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
      & & \tsplice{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}
\end{array}
\]

@@

    \node[main node] (rwxo) [below right of=rwx,yshift=-0.1cm] {$\rwxo$};
    \node[main node] (rxo) [below of=rwxo] {$\rxo$};
    \node[main node] (eo) [below of=rxo] {$\entero$};

@@

    \node[main node] (rwo) [below right of=rw,yshift=-0.1cm] {$\rwo$};
    \node[main node] (ro) [below of=rwo] {$\ro$};
    \node[main node] (0o) [below of=ro] {$\nopermo$};

    \path[every node/.style={font=\sffamily\small}]
    (rwxo) edge (rxo)
    (rxo) edge (eo)

@@

    (rwo) edge (ro)
    (ro) edge (0o)

    (rwx) edge[dashed] (rx)
    (rx) edge[dashed] (e)

    (rwo) edge (rwxo)
    (ro) edge (rxo)
    (0o) edge (eo)

    (rwo) edge (rw)
    (ro) edge (r)
    (0o) edge (0)

    (rw) edge (rwx)
    (r) edge[dashed] (rx)
    (0) edge[dashed] (e)

    (rwxo) edge (rwx)
    (rxo) edge[dashed] (rx)
    (eo) edge[dashed] (e);
  \end{tikzpicture}

@@

\subsection{Operational Semantics}
\subsubsection{Notes}
Genrally:
\begin{itemize}
\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
\item Linear capabilities are cleared when they move around in memory.
\end{itemize}

Source language:
\begin{itemize}
\item Variable length instructions that match the length of the compiled instructions
  \begin{itemize}
  \item Opaque capabilities do not hide the length of a program (especially an issue if we ever hope to have error recovery).
  \item This is needed for correctness. See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
  \end{itemize}
\item Programs have undefined behavior when passing ret/stk pointers.
\end{itemize}

@@

\[
  \updPcPerm{c} \defeq \dots
\]

@@

\[
  \isLinear{\cb} \defeq \dots
\]

\[
  \nonLinear{\cb} \defeq \dots
\]

@@

\[
  \opaquePerm{} \defeq \dots
\]

@@

\[
  \nonZero{c} \defeq \dots
\]

\subsubsection{Source Language}
\label{sec:source-language}
\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}

@@

\begin{align*}
  \tsem{\tgeta{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{\aaddr}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \text{ or } \\
                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} \tand\\
                                  \perm \not\in \opaquePerm{}
                                \end{array} \\
                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                              \end{cases}
\end{align*}

\begin{align*}
  \tsem{\sgetb{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{\baddr}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \text{ or } \\
                                  \quad \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} \tand\\
                                  \perm \not\in \opaquePerm{}
                                \end{array} \\
                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                              \end{cases}
\end{align*}

\begin{align*}
  \tsem{\sgete{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{\eaddr}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \text{ or } \\
                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} \tand\\
                                  \perm \not\in \opaquePerm{}
                                \end{array} \\
                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                              \end{cases}
\end{align*}

@@

\textbf{jmp and jnz}\\
Notice that the order of the updates of the configuration actually matters here. As it is below, a successful jump to $\pcreg$ results in a loop, but if the order of the updates to the configuration had been reversed, then the execution would fail as the $\pcreg$ register would contain 0.
\begin{align*}
  \tsem{\tjmp{r}} = & 
                      \begin{cases}
                        \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_r)}
                             & \nonLinear{\Phi(r)} \\
                        \Phi\updReg{\pcreg}{\Phi(r)} & \totherwise
                      \end{cases}
\end{align*}

>>>>

% Text
\newcommand{\tand}{\text{ and }}
\newcommand{\tor}{\text{ or }}
\newcommand{\totherwise}{\text{otherwise }}

@@

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{gettype}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
%\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
\newcommand{\sgetlin}[2]{\twoinstr{\src{get}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
\newcommand{\scseal}[3]{\threeinstr{\src{cseal}}{#1}{#2}{#3}}
\newcommand{\ssplice}[3]{\threeinstr{\src{splice}}{#1}{#2}{#3}}
\newcommand{\ssplit}[3]{\threeinstr{\src{split}}{#1}{#2}{#3}}

@@

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{gettype}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
%\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
\newcommand{\tgetlin}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}

% Three arguments
\newcommand{\tsplice}[3]{\threeinstr{\trg{splice}}{#1}{#2}{#3}}
\newcommand{\tsplit}[3]{\threeinstr{\trg{split}}{#1}{#2}{#3}}
\newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
\newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
\newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
\newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}
\newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}

@@

%%% Permissions
\newcommand{\plainperm}[1]{\mathrm{#1}}
%\newcommand{\rwlx}{\plainperm{rwlx}}
\newcommand{\rwx}{\plainperm{rwx}}
\newcommand{\rx}{\plainperm{rx}}
%\newcommand{\rwlxo}{\plainperm{rwlxo}}
%\newcommand{\rwlo}{\plainperm{rwlo}}
%\newcommand{\rwl}{\plainperm{rwl}}
%\newcommand{\rwxo}{\plainperm{rwxo}}
%\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
%\newcommand{\rxo}{\plainperm{rxo}}
%\newcommand{\ro}{\plainperm{ro}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\noperm}{\plainperm{0}}
%\newcommand{\nopermo}{\plainperm{0o}}
\newcommand{\enter}{\plainperm{e}}
%\newcommand{\entero}{\plainperm{eo}}

@@

\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
\newcommand{\encPerm}[1]{\plainfun{encocePermission}{#1}}
\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}

@@

\subsection{Syntax}
\label{sec:syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI). The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \tRegName \\
\trg{\rn} &\defbnf &  \trg{r} \mid n \\
\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
      & & \tsplice{\trg{r}}{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}{\trg{r}}
\end{array}
\]

@@

@@

    \path[every node/.style={font=\sffamily\small}]

@@

    (rwx) edge (rx)
    (rx) edge (e)

    (rw) edge (rwx)
    (r) edge (rx)
    (0) edge (e);
  \end{tikzpicture}

@@

\subsection{Operational Semantics}
\subsubsection{Notes}
Genrally:
\begin{itemize}
%\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
\item Linear capabilities are cleared when they move around in memory.
\end{itemize}

Source language:
\begin{itemize}
\item Variable length instructions that match the length of the compiled instructions
  \begin{itemize}
%  \item Opaque capabilities do not hide the length of a program (especially an issue if we ever hope to have error recovery).
  \item This is needed for correctness. See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
  \end{itemize}
\item Programs have undefined behavior when passing ret/stk pointers.
\end{itemize}

@@

\[
  \updPcPerm{c} \defeq
  \begin{cases}
    ((\rx,\lin),\baddr,\eaddr,\aaddr) & c = ((\enter,\lin),\baddr,\eaddr,\aaddr)\\
    c & \totherwise
  \end{cases}
\]

@@

\[
  \isLinear{\cb} \defeq
  \begin{cases}
    \top & 
    \arraycolsep=0pt
    \begin{array}[t]{l}
      \cb = ((\_,\linear),\_,\_,\_) \tor\\
      \cb = \seal{\linear,\_,\_} \tor\\
      \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
    \end{array}\\
    \src{\top} & 
    \arraycolsep=0pt
    \begin{array}[t]{l}
      \src{\cb = \stkptr{\_,\_,\_,\_}}\\
      \src{\cb = \retptrd{}}
    \end{array}\\
    \bot & \totherwise
  \end{cases}
\]

\[
  \nonLinear{\cb} \defeq \neg \isLinear{\cb}
\]

@@

@@

\[
  \nonZero{w} \defeq
  \begin{cases}
    \bot & w \in \ints \tand w = 0 \\
    \top & \totherwise
  \end{cases}
\]

\subsubsection{The Operational Semantics}
We have unified the two languages in the below definitions. Everything written in black is common for both source and target language. Everything written in \src{blue} is specific to the source language.

\noindent\textbf{fail and halt}\\
\begin{align*}
  \sem{\tfail}(\Phi) = & \; \failed \\
  \sem{\thalt}(\Phi) = & \; (\halted, \Phi.\mem)
\end{align*}

\noindent\textbf{jmp and jnz}\\
\begin{align*}
  \sem{\tjmp{r}}(\Phi) = &  
                     \begin{cases}
                       \Phi\updReg{r}{0}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & \isLinear{\Phi(r)} \\
                       \Phi\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & \totherwise
                     \end{cases}
\end{align*}

\begin{align*}
  \sem{\tjnz{r_1}{r_2}}(\Phi) = &       
                             \begin{cases}
                               \Phi\updReg{\pcreg}{r_1} & \text{if $\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
                               \Phi\updReg{r_1}{0}\updReg{\pcreg}{\Phi(r_1)}
                                               & \arraycolsep=0pt                               
                                               \begin{array}[t]{l}
                                                 \isLinear{\Phi(r_1)} \tand
                                                 \nonZero{\Phi{r_2}}
                                               \end{array}
\\
                               \updPcAddr{\Phi} & \totherwise
                             \end{cases}
\end{align*}

\noindent\textbf{gettype}\\
In the definitions of the semantics below, we use a function $\encType{} : \Word \rightarrow \ints$. This is an encoding function for which the specific implementation does not matter. As the words of the two machines differ, we really need two functions which we call $\encType{}_{\var{src}}$ and $\encType{}_{\var{trg}}$. These two functions need to be related in the following way:
\lau{14-09-2017: Insert the relation. My impression is that the the words specific to the source machine should have the same "type" as their implementaion at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation). }
\begin{align*}
  \sem{\tisptr{r_1}{r_2}}(\Phi) = & \; \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
\end{align*}

\noindent\textbf{geta, getb, gete, getp, and getl}\\
We assume function to encode and decode permissions as well as a function to encode and decode linearity. The functions are used implicitely when a permission or linearity is used in a place where they need to be a word.
\lau{14-09-2017: Write more about this (types and assumptions)}
\begin{align*}
  \sem{\tgeta{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\aaddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgetb{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\baddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgete{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\eaddr}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgetp{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\perm}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\perm,\_),\_,\_,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\_,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\begin{align*}
  \sem{\tgetlin{r_1}{r_2}}(\Phi) = & 
                                   \begin{cases}
                                     \updPcAddr{\Phi\updReg{r_1}{\lin}} & 
                                     \arraycolsep=0pt
                                     \begin{array}[t]{l}
                                       \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \\
                                     \end{array} \\
                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                                   \end{cases}
\end{align*}

\noindent\textbf{move}\\
\begin{align*}
  \sem{\tmove{r_1}{r_2}}(\Phi) = & 
                              \begin{cases}
                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
                              \end{cases}
\end{align*}

\noindent\textbf{call}\\
\begin{align*}
  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & \dots
\end{align*}

\noindent\textbf{cseal}\\
\begin{align*}
  \sem{\tcseal{r_1}{r_2}{r_3}}(\Phi) = &
                                  \begin{cases}
                                    \Phi\updReg{r_1}{\vsc} &
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) \in \SealableCaps \tand \\
                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr} \tand \\
                                      \sigma_\baddr \leq \sigma_\eaddr \tand \\ 
                                      \vsc = \sealed{c,\sigma}
                                    \end{array}
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

\textbf{split}\\
\lau{14-09-2017: Do we want split to work on seals (not sealed capabilities, but seals?)}
\begin{align*}
  \tsem{\tsplit{r_1}{r_2}{r_3}} = &
                               \begin{cases}
                                 \Phi\updReg{r_1}{\dots} &
                                 \arraycolsep=0pt
                                 \begin{array}[t]{l}
                                   \dots
                                 \end{array}\\
                                 \failed & \totherwise
                               \end{cases}
\end{align*}

\subsubsection{Source Language}
\label{sec:source-language}
\lau{14-09-2017: This section will be obsolete when the unifyed definitions are done.}
\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}

@@

\begin{align*}
  \tsem{\tgeta{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{\aaddr}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \text{ or } \\
                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} \tand\\
                                \end{array} \\
                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                              \end{cases}
\end{align*}

\begin{align*}
  \tsem{\sgetb{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{\baddr}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \text{ or } \\
                                  \quad \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} \tand\\
                                \end{array} \\
                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                              \end{cases}
\end{align*}

\begin{align*}
  \tsem{\sgete{r_1}{r_2}} = &
                              \begin{cases}
                                \updPcAddr{\Phi\update{\reg.r_1}{\eaddr}} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \text{ or } \\
                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} \tand\\
                                \end{array} \\
                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
                              \end{cases}
\end{align*}

@@

\textbf{jmp and jnz}\\
Notice that the order of the updates of the configuration actually matters here. As it is below, a successful jump to $\pcreg$ results in a loop, but if the order of the updates to the configuration had been reversed, then the execution would fail as the $\pcreg$ register would contain 0.
\begin{align*}
  \tsem{\tjmp{r}} = & 
                      \begin{cases}
                        \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_r)}
                             & \isLinear{\Phi(r)} \\
                        \Phi\updReg{\pcreg}{\Phi(r)} & \totherwise
                      \end{cases}
\end{align*}
